\documentclass[11pt,twoside,openright]{report}
\usepackage[a4paper,width=155mm, inner=25mm, top=30mm,bottom=40mm, marginparsep=0mm]{geometry}
\usepackage[utf8]{inputenc} 
\usepackage[T1]{fontenc} 
\usepackage[czech]{babel} 
\usepackage{graphicx}
\usepackage{emptypage} % maže číslování na prázdných stranách
\usepackage[numbers,sort&compress]{natbib}
\usepackage{formatting,amsfonts,amsmath,amssymb}
\usepackage{booktabs,mathtools}
\usepackage{lipsum,url}
\usepackage[unicode]{hyperref}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{array}
\usepackage{float}
\usepackage{mwe}
\usepackage{appendix}
\usepackage{pdfpages}
\usepackage{etoolbox}
\usepackage[nottoc]{tocbibind}
\usepackage{tikz}
\usepackage{enumerate}
%\usepackage{enumitem}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{csquotes}
\usetikzlibrary{shapes, arrows.meta}
\usetikzlibrary{backgrounds,calc}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\BeforeBeginEnvironment{appendices}{\gdef\thechapter{\ }}
\AtBeginEnvironment{appendices}{\renewcommand\thesection{\Alph{section}}}

% \cmidrule have problem with Czech
\usepackage{regexpatch}
\makeatletter
% Change the `-` delimiter to an active character
\xpatchparametertext\@@@cmidrule{-}{\cA-}{}{}
\xpatchparametertext\@cline{-}{\cA-}{}{}
\makeatother

%\usepackage[nottoc,numbib]{tocbibind}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{subfig}
\usepackage{fancybox}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
\makeatletter
\renewcommand{\ALG@name}{Algoritmus}
\makeatother
\pagestyle{fancy}
%\renewcommand{\chaptermark}[1]{%
%  \markboth{\MakeUppercase{\thechapter\ \quad  #1}}{}% 
%}
\renewcommand{\chaptermark}[1]{%
  \markboth{\MakeUppercase{#1}}{}% 
}
\fancyfoot{}{}{}
\fancyhead[LO,LE]{}
\fancyhead[RO]{\leftmark}
\fancyhead[RE]{}
\fancyfoot[RO]{\vspace*{0.8\baselineskip}\thepage}
\fancyfoot[LE]{\vspace*{0.8\baselineskip}\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\fancypagestyle{plain}{%
\fancyhead[RO]{}
\fancyhead[LO,LE]{}
\fancyfoot[RO]{\vspace*{0.8\baselineskip}\thepage}
\fancyfoot[LE]{\vspace*{0.8\baselineskip}\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
}

\linespread{1.25}
\begin{document}
\pagenumbering{gobble}

\input{VNITRNOSTI/main_info}
% TAK AT TO JDE OD RUKY

\input{VNITRNOSTI/uvod}
\input{VNITRNOSTI/adaptivni_filtry}
\input{VNITRNOSTI/nd_metody}

\input{VNITRNOSTI/gpd}
%\input{VNITRNOSTI/cile}
\input{VNITRNOSTI/ese}
\input{VNITRNOSTI/vysledky}
\input{VNITRNOSTI/zaver}


\renewcommand{\bibname}{Publikace autora}
\begin{thebibliography}{A}
\bibitem{ese_mdpi}VRBA, Jan; MAREŠ, Jan. Introduction to Extreme Seeking Entropy. Entropy, 2020, 22.1: 93.
\bibitem{ijcnn}BUKOVSKÝ, Ivo, CEJNEK, Matouš, VRBA, Jan, HOMMA, Noryasu. Study of Learning Entropy for Onset Detection of Epileptic Seizures in EEG Time Series. \textit{The 2016 International Joint Conference on Neural Networks (IJCNN 2016)}. IEEE WCCI 2016, Vancouver, 2016. ISBN 978-1-5090-0620-5
\bibitem{appel1}VRBA, Jan. Adaptive Novelty Detection with Generalized Extreme Value Distribution. In: 2018 International Conference on Applied Electronics (AE). IEEE, 2018. p. 1-4.
\bibitem{cyril}OSWALD, Cyril, CEJNEK, Matouš, VRBA, Jan, BUKOVSKÝ, Ivo. Novelty Detection in System Monitoring and Control with HONU. In: ZHANG, Ming. \textit{ Aplied artificial higher order neural networks for control and recognition}. Hershey, PA: Information Science Reference, an imprint of IGI Global, 2016. ISBN 9781522500643.
\bibitem{roboti}VRBA, Jan, et al. An Automated Platform for Microrobot Manipulation. In: International Workshop on Soft Computing Models in Industrial and Environmental Applications. Springer, Cham, 2020. p. 255-265.
\bibitem{bila1}BÍLA, Jiří; VRBA, Jan. The Detection and Interpretation of Emergent Situations in ECG Signals. In: International Conference on Soft Computing-MENDEL. Springer, Cham, 2016. p. 264-275.
\bibitem{ijcnn3}MOJZES, Matej, et al. Feature selection via competitive levy flights. In: 2016 International Joint Conference on Neural Networks (IJCNN). IEEE, 2016. p. 3731-3736.
\bibitem{bila2}BILA, Jiri; NOVAK, Martin; VRBA, Jan. Detection of emergent situations in complex systems represented by algebras of transformations. In: MATEC Web of Conferences. EDP Sciences, 2016. p. 02035.
\bibitem{artep}VRBA, Jan. Využití fuzzy systémů a algoritmu learning entropy pro detekci změn stavů bioprocesu. In: \textit{Automatizacia a riadenie v teorii a praxi ARTEP 2017}. 11. ročník konferencie odbornikov z univerzit, vysokých škol a praxe. Technická univerzita Košice, 2017. ISBN 978-80-553-3075-4.
\bibitem{asr}VRBA, Jan, ŠKOLÁKOVÁ, Andrea. Adaptivní metoda detekce novosti, XLIII. In: \textit{ Seminář ASŘ 2018}. VŠB TU Ostrava, 2018.
\bibitem{artep2}BUKOVSKÝ, Ivo, OSWALD, Cyril, VRBA, Jan. Případová studie použití entropie učení pro adaptivní detekci při řízení spalování tuhých paliv. In: \textit{Automatizácia a riadenie v teórii a praxi 2015}. Technická univerzita Košice, 2015. ISBN 978-80-553-1968-1.
\bibitem{appel2}VRBA, Jan; MAREŠ, Jan. Computational Performance of the Parameters Estimation in Extreme Seeking Entropy Algorithm. In: 2020 International Conference on Applied Electronics (AE). IEEE, 2020. p. 1-4.
\bibitem{appel3}VRBA, Jan; MAREŠ, Jan. ROC Analysis of Extreme Seeking Entropy for Trend Change Detection. In: 2020 International Conference on Applied Electronics (AE). IEEE, 2020. p. 1-4.
\end{thebibliography}


\renewcommand{\bibname}{Literatura}
\begin{thebibliography}{L}

\bibitem{fermentace}MAREŠ, Jan, et al. Process state classification of fed-batch fermentation based on process variables analysis. \textit{Biochemical Engineering Journal}, 2016, 112: 178-185.
\bibitem{rls_mit}ROWELL, D. 2.161 Signal Processing: Continuous and Discrete, Fall 2008. 2008.
\bibitem{rls_kalman}MANDIC, Danilo P.; KANNA, Sithan; CONSTANTINIDES, Anthony G. On the intrinsic relationship between the least mean square and Kalman filters [Lecture Notes]. \textit{IEEE Signal Processing Magazine}, 2015, 32.6: 117-122.
\bibitem{proakis}PROAKIS, John G. a Dimitris G. MANOLAKIS. \textit{Digital signal processing: principles, algorithms, and applications}. Pearson Education India, 2001.
\bibitem{eeg}OOSTENVELD, Robert; PRAAMSTRA, Peter. The five percent electrode system for high-resolution EEG and ERP measurements. \textit{Clinical neurophysiology}, 2001, 112.4: 713-719.
\bibitem{haykin}HAYKIN, Simon S. \textit{Adaptive filter theory}. Pearson Education India, 2008.
\bibitem{dirac}STRICHARTZ, Robert S. \textit{A guide to distribution theory and Fourier transforms}. World Scientific Publishing Company, 2003.
\bibitem{kvantizace}GERSHO, Allen; GRAY, Robert M. \textit{Vector quantization and signal compression}. Springer Science \& Business Media, 2012.
\bibitem{ivoLE2}BUKOVSKY, Ivo; KINSNER, Witold; HOMMA, Noriyasu. Learning Entropy as a Learning-Based Information Concept. \textit{Entropy}, 2019, 21.2: 166.
\bibitem{ivoLE1}BUKOVSKY, Ivo. Learning entropy: Multiscale measure for incremental learning. \textit{Entropy}, 2013, 15.10: 4159-4187.
\bibitem{elbnd1}CEJNEK, Matous; BUKOVSKY, Ivo. Concept drift robust adaptive novelty detection for data streams. \textit{Neurocomputing}, 2018, 309: 46-53.
\bibitem{elbnd2} CEJNEK, Matous; BUKOVSKY, Ivo. Influence of type and level of noise on the performance of an adaptive novelty detector. In: \textit{2017 IEEE 16th International Conference on Cognitive Informatics \& Cognitive Computing (ICCI* CC)}. IEEE, 2017. p. 373-377.
\bibitem{elbnd3}CEJNEK, Matous; BUKOVSKY, Ivo; VYSATA, Oldrich. Adaptive classification of EEG for dementia diagnosis. In: \textit{2015 International Workshop on Computational Intelligence for Multimedia Understanding (IWCIM)}. IEEE, 2015. p. 1-5.
\bibitem{fault}GERTLER, Janos. \textit{Fault detection and diagnosis in engineering systems}. CRC press, 1998.

\bibitem{data_streams}YU, Kangqing, et al. Real-time Outlier Detection over Streaming Data. In: \textit{2019 IEEE SmartWorld, Ubiquitous Intelligence \& Computing, Advanced \& Trusted Computing, Scalable Computing \& Communications, Cloud \& Big Data Computing, Internet of People and Smart City Innovation (SmartWorld/SCALCOM/UIC/ATC/CBDCom/IOP/SCI)}. IEEE, 2019. p. 125-132.

\bibitem{surveilance}
RAMEZANI, Ramin; ANGELOV, Plamen; ZHOU, Xiaowei. A fast approach to novelty detection in video streams using recursive density estimation. In: \textit{2008 4th International IEEE Conference Intelligent Systems. IEEE, 2008}. p. 14-2-14-7.
\bibitem{robotics_marslan}
MARSLAND, Stephen; NEHMZOW, Ulrich; SHAPIRO, Jonathan. On-line novelty detection for autonomous mobile robots. \textit{Robotics and Autonomous Systems}, 2005, 51.2-3: 191-206.

\bibitem{robotics}
NEHMZOW, Ulrich, et al. Novelty detection as an intrinsic motivation for cumulative learning robots. In: \textit{Intrinsically Motivated Learning in Natural and Artificial Systems}. Springer, Berlin, Heidelberg, 2013. p. 185-207.
\bibitem{scarrot2012review}SCARROTT, Carl; MACDONALD, Anna. A review of extreme value threshold es-timation and uncertainty quantification. \textit{REVSTAT–Statistical Journal}, 2012, 10.1: 33-60.
\bibitem{mackey}MACKEY, Michael C.; GLASS, Leon. Oscillation and chaos in physiological control systems. \textit{Science}, 1977, 197.4300: 287-289.
\bibitem{dead}SPANGENBERG, Mariana, et al. Detection of variance changes and mean value jumps in measurement noise for multipath mitigation in urban navigation. \textit{Navigation}, 2010, 57.1: 35-52.
\bibitem{diagnosis}MAURYA, Mano Ram; RENGASWAMY, Raghunathan; VENKATASUBRAMANIAN, Venkat. Fault diagnosis using dynamic trend analysis: A review and recent developments. \textit{Engineering Applications of artificial intelligence}, 2007, 20.2: 133-146.
\bibitem{stepchange}L'ECUYER, Pierre. History of uniform random number generation. In: \textit{2017 Winter Simulation Conference (WSC)}. IEEE, 2017. p. 202-230.
\bibitem{roc_orig}EGAN, James P. \textit{Signal detection theory and ROC-analysis}. Academic press, 1975.

\bibitem{roc_bible}FAWCETT, Tom. An introduction to ROC analysis. \textit{Pattern recognition letters}, 2006, 27.8: 861-874.

\bibitem{mom_orig}HOSKING, Jonathan RM; WALLIS, James R. Parameter and quantile estimation for the generalized Pareto distribution. \textit{Technometrics}, 1987, 29.3: 339-349.
\bibitem{DuMouchel}DUMOUCHEL, William H. Estimating the stable index $\alpha$ in order to measure tail thickness: a critique. \textit{The Annals of Statistics}, 1983, 1019-1031.
\bibitem{history}HALD, Anders. \textit{A History of Mathematical Statistics from 1750 to 1930}. New York: Wiley, 1998.

\bibitem{stat_bible}MONTGOMERY, Douglas C.; RUNGER, George C. \textit{Applied Statistics and Probability for Engineers}. Wiley, 2007.

\bibitem{grim}GRIMSHAW, Scott D. Computing maximum likelihood estimates for the generalized Pareto distribution. \textit{Technometrics}, 1993, 35.2: 185-191.
\bibitem{SmithML}SMITH, Richard L. Maximum likelihood estimation in a class of nonregular cases. \textit{Biometrika}, 1985, 72.1: 67-90.
\bibitem{Luceno}LUCENO, Alberto. Fitting the generalized Pareto distribution to data using maximum goodness-of-fit estimators. \textit{Computational Statistics \& Data Analysis}, 2006, 51.2: 904-917.
\bibitem{zhang1}ZHANG, Jin. Likelihood moment estimation for the generalized Pareto distribution. \textit{Australian \& New Zealand Journal of Statistics}, 2007, 49.1: 69-77.
\bibitem{zhao}ZHAO, Xu, et al. A New Parameter Estimator for the Generalized Pareto Distribution under the Peaks over Threshold Framework. \textit{Mathematics}, 2019, 7.5: 406.
\bibitem{park}PARK, Myung Hyun; KIM, Joseph HT. Estimating extreme tail risk measures with generalized Pareto distribution. \textit{Computational Statistics \& Data Analysis}, 2016, 98: 91-104.
\bibitem{gpd_est}DE ZEA BERMUDEZ, P.; KOTZ, Samuel. Parameter estimation of the generalized Pareto distribution—Part I. \textit{Journal of Statistical Planning and Inference}, 2010, 140.6: 1353-1373.
\bibitem{miljkovic}MILJKOVIĆ, Dubravko. Review of novelty detection methods. In: \textit{The 33rd International Convention MIPRO}. IEEE, 2010. p. 593-598.
\bibitem{zimek}ZIMEK, Arthur; FILZMOSER, Peter. There and back again: Outlier detection between statistical reasoning and data mining algorithms. \textit{Wiley Interdisciplinary Reviews: Data Mining and Knowledge Discovery}, 2018, 8.6: e1280.

\bibitem{markou1}MARKOU, Markos; SINGH, Sameer. Novelty detection: a review—part 1: statistical approaches. \textit{Signal processing}, 2003, 83.12: 2481-2497.
\bibitem{markou2}MARKOU, Markos; SINGH, Sameer. Novelty detection: a review—part 2:: neural network based approaches. \textit{Signal processing}, 2003, 83.12: 2499-2521.
\bibitem{pimentel}PIMENTEL, Marco AF, et al. A review of novelty detection. \textit{Signal Processing}, 2014, 99: 215-249.
\bibitem{set_member}GOLLAMUDI, Sridhar, et al. Set-membership filtering and a set-membership normalized LMS algorithm with an adaptive step size. \textit{IEEE Signal Processing Letters}, 1998, 5.5: 111-114.
\bibitem{diniz1}DINIZ, Paulo SR; YAZDANPANAH, Hamed. Data censoring with set-membership algorithms. In: \textit{2017 IEEE Global Conference on Signal and Information Processing (GlobalSIP)}. IEEE, 2017. p. 121-125.
\bibitem{set_member_robot}ZHOU, Bo, et al. Multi-sensor fusion robust localization for indoor mobile robots based on a set-membership estimator. In: \textit{2015 IEEE International Conference on Cyber Technology in Automation, Control, and Intelligent Systems (CYBER)}. IEEE, 2015. p. 157-162.
\bibitem{set_member_robot_voda}JAULIN, Luc. Robust set-membership state estimation; application to underwater robotics. \textit{Automatica}, 2009, 45.1: 202-206.
\bibitem{set_member_robot_chyba}HUANG, Jian; WANG, Yuan; FUKUDA, Toshio. Set-membership-based fault detection and isolation for robotic assembly of electrical connectors. \textit{IEEE Transactions on Automation Science and Engineering}, 2016, 15.1: 160-171.
\bibitem{diniz0}DINIZ, Paulo Sergio Ramirez. \textit{Adaptive filtering: algorithms and practical implementation}. Springer, 2008.
\bibitem{auto1}NICOLAU, Miguel, et al. A hybrid autoencoder and density estimation model for anomaly detection. In: \textit{International Conference on Parallel Problem Solving from Nature}. Springer, Cham, 2016. p. 717-726.
\bibitem{auto2}AN, Jinwon; CHO, Sungzoon. Variational autoencoder based anomaly detection using reconstruction probability. \textit{Special Lecture on IE}, 2015, 2.1: 1-18.
\bibitem{auto3}MARCHI, Erik, Fabio VESPERINI, Florian EYBEN, Stefano SQUARTINI a Bjorn SCHULLER. A novel approach for automatic acoustic novelty detection using a denoising autoencoder with bidirectional LSTM neural networks. In: \textit{2015 IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP)} 2015, p. 1996-2000.
\bibitem{auto4}FENG, Qingsong, et al. Anomaly detection of spectrum in wireless communication via deep auto-encoders. \textit{The Journal of Supercomputing}, 2017, 73.7: 3161-3178.
\bibitem{auto5}DOMINGUES, Rémi, et al. Deep Gaussian Process autoencoders for novelty detection. \textit{Machine Learning}, 2018, 107.8-10: 1363-1383.
\bibitem{auto6}BHATTACHARJEE, Supritam; MANDAL, Devraj; BISWAS, Soma. Autoencoder based novelty detection for generalized zero shot learning. In: \textit{2019 IEEE International Conference on Image Processing (ICIP)}. IEEE, 2019. p. 3646-3650.
\bibitem{auto7}KIEU, Tung, et al. Outlier Detection for Time Series with Recurrent Autoencoder Ensembles. In: \textit{IJCAI}. 2019. p. 2725-2732.
\bibitem{auto8}MARCHI, Erik, et al. Deep recurrent neural network-based autoencoders for acoustic novelty detection. \textit{Computational intelligence and neuroscience}, 2017, 2017.
\bibitem{auto9}KAWACHI, Yuta; KOIZUMI, Yuma; HARADA, Noboru. Complementary set variational autoencoder for supervised anomaly detection. In: \textit{2018 IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP)}. IEEE, 2018. p. 2366-2370.
\bibitem{auto10}WANG, Xuhong, et al. adVAE: A self-adversarial variational autoencoder with Gaussian anomaly prior knowledge for anomaly detection. \textit{Knowledge-Based Systems}, 2020, 190: 105187.
\bibitem{gmm1}REYNOLDS, Douglas A. Gaussian Mixture Models. \textit{Encyclopedia of biometrics}, 2009, 741.
\bibitem{gmm2}HEYNS, Theo; HEYNS, Philippus Stephanus; DE VILLIERS, Johan Pieter. Combining synchronous averaging with a Gaussian mixture model novelty detection scheme for vibration-based condition monitoring of a gearbox. \textit{Mechanical Systems and Signal Processing}, 2012, 32: 200-215.
\bibitem{gmm3}BISHOP, Christopher M. \textit{Pattern recognition and machine learning}. Springer, 2006.
\bibitem{gmm4}DUDA, Richard O.; HART, Peter E.; STORK, David G. \textit{Pattern classification}. John Wiley \& Sons, 2012.
\bibitem{gmm5}DOAN, AnHai; HALEVY, Alon; IVES, Zachary. \textit{Principles of data integration}. Elsevier, 2012.
\bibitem{gmm6}ZHANG, Yu, et al. Detection of emerging faults on industrial gas turbines using extended Gaussian mixture models. \textit{International Journal of Rotating Machinery}, 2017, 2017.
\bibitem{gmm7}NASIOS, Nikolaos; BORS, Adrian G. Variational learning for Gaussian mixture models. \textit{IEEE Transactions on Systems, Man, and Cybernetics, Part B (Cybernetics)}, 2006, 36.4: 849-862.
\bibitem{gmm8}MIYAZAKI, Bungo, et al. Change detection of orders in stock markets using a Gaussian mixture model. \textit{Intelligent Systems in Accounting, Finance and Management}, 2014, 21.3: 169-191.
\bibitem{gmm9}DEMPSTER, Arthur P.; LAIRD, Nan M.; RUBIN, Donald B. Maximum likelihood from incomplete data via the EM algorithm. \textit{Journal of the Royal Statistical Society: Series B (Methodological)}, 1977, 39.1: 1-22.
\bibitem{gmm10}WATANABE, Hidenori; MURAMATSU, Shogo; KIKUCHI, Hisakazu. Interval calculation of EM algorithm for GMM parameter estimation. In: \textit{Proceedings of 2010 IEEE International Symposium on Circuits and Systems}. IEEE, 2010. p. 2686-2689.
\bibitem{gmm11}BOUGUILA, Nizar a Wentao FAN, ed. \textit{Mixture Models and Applications} Springer International Publishing, 2020.
\bibitem{ferreira2003optimising}FERREIRA, A.; DE HAAN, Laurens; PENG, Liang. On optimising the estimation of high quantiles of a probability distribution. \textit{Statistics}, 2003, 37.5: 401-434.
\bibitem{loretan1994testing}LORETAN, Mico; PHILLIPS, Peter CB. Testing the covariance stationarity of heavy-tailed time series: An overview of the theory with applications to several financial datasets. \textit{Journal of empirical finance}, 1994, 1.2: 211-248.
\bibitem{scarrott2012review}SCARROTT, Carl; MACDONALD, Anna. A review of extreme value threshold estimation and uncertainty quantification. \textit{REVSTAT–Statistical Journal}, 2012, 10.1: 33-60.
\bibitem{gpd1}FORBES, Catherine, et al. \textit{Statistical distributions}. John Wiley \& Sons, 2011.
\bibitem{gpd2}COLES, Stuart, et al. \textit{An introduction to statistical modeling of extreme values}. London: Springer, 2001.
\bibitem{gpd3}LEE, Hyoung-joo; ROBERTS, Stephen J. On-line novelty detection using the kalman filter and extreme value theory. In: \textit{2008 19th International Conference on Pattern Recognition}. IEEE, 2008. p. 1-4.
\bibitem{gpd4}LIMPERT, Eckhard; STAHEL, Werner A. Problems with using the normal distribution–and ways to improve quality and efficiency of data analysis. \textit{PLoS One}, 2011, 6.7: e21403.
\bibitem{grubbs}GRUBBS, Frank E. Procedures for detecting outlying observations in samples. \textit{Technometrics}, 1969, 11.1: 1-21.
\bibitem{gpd5}PICKANDS III, James, et al. Statistical inference using extreme order statistics. \textit{The Annals of Statistics}, 1975, 3.1: 119-131.
\bibitem{gpd6}BALKEMA, August A.; DE HAAN, Laurens. Residual life time at great age. \textit{The Annals of probability}, 1974, 792-804.
\bibitem{evt1}ROBERTS, Stephen J. Novelty detection using extreme value statistics. \textit{IEE Proceedings-Vision, Image and Signal Processing}, 1999, 146.3: 124-129.
\bibitem{evt2}FISHER, Ronald Aylmer; TIPPETT, Leonard Henry Caleb. Limiting forms of the frequency distribution of the largest or smallest member of a sample. In: \textit{Mathematical Proceedings of the Cambridge Philosophical Society}. Cambridge University Press, 1928. p. 180-190.
\bibitem{evt3}CLIFTON, David Andrew; HUGUENY, Samuel; TARASSENKO, Lionel. Novelty detection with multivariate extreme value statistics. \textit{Journal of signal processing systems}, 2011, 65.3: 371-389.
\bibitem{maha}MAHALANOBIS, Prasanta Chandra. \textit{On the generalized distance in statistics}. National Institute of Science of India, 1936.
\bibitem{kalman}GREWAL, Mohinder S.; ANDREWS, Angus P. \textit{Kalman filtering: Theory and Practice with MATLAB}. John Wiley \& Sons, 2014.
\bibitem{evt4}LEE, Hyoung-joo; ROBERTS, Stephen J. On-line novelty detection using the kalman filter and extreme value theory. In: \textit{2008 19th International Conference on Pattern Recognition}. IEEE, 2008. p. 1-4.
\bibitem{evt5}CLIFTON, David A., et al. Extending the generalised pareto distribution for novelty detection in high-dimensional spaces. \textit{Journal of signal processing systems}, 2014, 74.3: 323-339.
\bibitem{evt6}SIFFER, Alban, et al. Anomaly detection in streams with extreme value theory. In: \textit{Proceedings of the 23rd ACM SIGKDD International Conference on Knowledge Discovery and Data Mining}. 2017. p. 1067-1075.
\bibitem{evt7}HUANG, Yan, et al. Generalized Pareto Model Based on Particle Swarm Optimization for Anomaly Detection. \textit{IEEE Access}, 2019, 7: 176329-176338.
\bibitem{pso}KENNEDY, James; EBERHART, Russell. Particle swarm optimization. In: \textit{Proceedings of ICNN'95-International Conference on Neural Networks}. IEEE, 1995. p. 1942-1948.
\bibitem{evt8}LUCA, Stijn; CLIFTON, David A.; VANRUMSTE, Bart. One-class classification of point patterns of extremes. \textit{The Journal of Machine Learning Research}, 2016, 17.1: 6581-6601.
\bibitem{svm1}BOSER, Bernhard E.; GUYON, Isabelle M.; VAPNIK, Vladimir N. A training algorithm for optimal margin classifiers. In: \textit{Proceedings of the fifth annual workshop on Computational learning theory}. 1992. p. 144-152.
\bibitem{svm2}CORTES, Corinna; VAPNIK, Vladimir. Support-vector networks. \textit{Machine learning}, 1995, 20.3: 273-297.
\bibitem{perceptron}GUPTA, Madan; JIN, Liang; HOMMA, Noriyasu. \textit{Static and dynamic neural networks: from fundamentals to advanced theory}. John Wiley \& Sons, 2004.
\bibitem{ml1}MOHRI, Mehryar; ROSTAMIZADEH, Afshin; TALWALKAR, Ameet. \textit{Foundations of machine learning}. MIT press, 2018.
\bibitem{ml2}PEDREGOSA, Fabian, et al. Scikit-learn: Machine learning in Python. \textit{The Journal of machine Learning research}, 2011, 12: 2825-2830.
\bibitem{svm3}MANNING, Christopher D.; SCHÜTZE, Hinrich; RAGHAVAN, Prabhakar. \textit{Introduction to information retrieval}. Cambridge university press, 2008.
\bibitem{svm4}CRISTIANINI, Nello, et al. \textit{An introduction to support vector machines and other kernel-based learning methods}. Cambridge university press, 2000.
\bibitem{kernel}AISERMAN, M. A.; BRAVERMAN, E. M.; ROZONOER, L. I. Theoretical foundations of the potential function method in pattern recognition. \textit{Avtomat. i Telemeh}, 1964, 25.6: 917-936.
\bibitem{svm5}SCHÖLKOPF, Bernhard, et al. Support vector method for novelty detection. In: \textit{Advances in neural information processing systems}. 2000. p. 582-588.
\bibitem{svm6}YAN, Ke; JI, Zhiwei; SHEN, Wen. Online fault detection methods for chillers combining extended kalman filter and recursive one-class SVM. \textit{Neurocomputing}, 2017, 228: 205-212.
\bibitem{svm7}SADOOGHI, Mohammad Saleh; KHADEM, Siamak Esmaeilzadeh. Improving one class support vector machine novelty detection scheme using nonlinear features. \textit{Pattern Recognition}, 2018, 83: 14-33.
\bibitem{pe}BANDT, Christoph; POMPE, Bernd. Permutation entropy: a natural complexity measure for time series. \textit{Physical review letters}, 2002, 88.17: 174102.
\bibitem{se}RICHMAN, Joshua S.; MOORMAN, J. Randall. Physiological time-series analysis using approximate entropy and sample entropy. \textit{American Journal of Physiology-Heart and Circulatory Physiology}, 2000, 278.6: H2039-H2049.
\bibitem{ae}PINCUS, Steven M. Approximate entropy as a measure of system complexity. \textit{Proceedings of the National Academy of Sciences}, 1991, 88.6: 2297-2301.
\bibitem{svm8}SOLARZ, A., et al. Automated novelty detection in the WISE survey with one-class support vector machines. \textot{Astronomy \& Astrophysics}, 2017, 606: A39.
\bibitem{somS}YIN, Hujun. Nonlinear dimensionality reduction and data visualization: a review. \textit{International Journal of Automation and Computing}, 2007, 4.3: 294-303.
\bibitem{somS2}AGUAYO, Leonardo; BARRETO, Guilherme A. Novelty detection in time series using self-organizing neural networks: a comprehensive evaluation. \textit{Neural Processing Letters}, 2018, 47.2: 717-744.
\bibitem{som0}KOHONEN, Teuvo. The self-organizing map. \textit{Proceedings of the IEEE}, 1990, 78.9: 1464-1480.
\bibitem{dynamic_neural}TAYLOR, John G. Neural ‘bubble’dynamics in two dimensions: foundations. \textit{Biological cybernetics}, 1999, 80.6: 393-409.

\bibitem{som1}BERNARD, Yann; HUEBER, Nicolas; GIRAU, Bernard. Novelty detection with self-organizing maps for autonomous extraction of salient tracking features. In: \textit{International Workshop on Self-Organizing Maps}. Springer, Cham, 2019. p. 100-109.

\bibitem{sal}CHI, Dongxiang. Self-organizing map-based color image segmentation with k-means clustering and saliency map. \textit{International Scholarly Research Notices, 2011}, 2011.

\bibitem{som2}RATRE, Avinash; PANKAJAKSHAN, Vinod. Tucker visual search-based hybrid tracking model and fractional Kohonen self-organizing map for anomaly localization and detection in surveillance videos. \textit{The Imaging Science Journal}, 2018, 66.4: 195-210.


\bibitem{som3}OLSZEWSKI, Dominik. Fraud detection using self-organizing map visualizing the user profiles. \textit{Knowledge-Based Systems}, 2014, 70: 324-334.

\bibitem{som4}RIGAMONTI, Marco, et al. Identification of the degradation state for condition-based maintenance of insulated gate bipolar transistors: A self-organizing map approach. \textit{Microelectronics Reliability}, 2016, 60: 48-61.
\bibitem{som5}MATIĆ, Frano; KALINIĆ, Hrvoje; VILIBIĆ, Ivica. Interpreting Self-Organizing Map errors in the classification of ocean patterns. \textit{Computers \& geosciences}, 2018, 119: 9-17.

\bibitem{som6}LI, Menglu; KASHEF, Rasha; IBRAHIM, Ahmed. Multi-Level Clustering-Based Outlier’s Detection (MCOD) Using Self-Organizing Maps. \textit{Big Data and Cognitive Computing}, 2020, 4.4: 24.
\bibitem{klastry}GAN, Guojun; MA, Chaoqun; WU, Jianhong. Data clustering: theory, algorithms, and applications. \textit{Society for Industrial and Applied Mathematics}, 2007.
\bibitem{som7}NAM, Tran Manh, et al. Self-organizing map-based approaches in DDoS flooding detection using SDN. In: \textit{2018 International Conference on Information Networking (ICOIN)}. IEEE, 2018. p. 249-254.
\bibitem{som8}ELMOUGY, Samir, et al. A parameter based growing ensemble of self-organizing maps for outlier detection in healthcare. \textit{Cluster Computing}, 2019, 22.1: 2437-2460.



\bibitem{cnn1}SIMONYAN, Karen; ZISSERMAN, Andrew. Very deep convolutional networks for large-scale image recognition. \textit{arXiv preprint arXiv:1409.1556}, 2014.
\bibitem{deep}GOODFELLOW, Ian, et al. \textit{Deep learning}. Cambridge: MIT press, 2016.
\bibitem{cnn2}TANG, Zhiyi, et al. Convolutional neural network‐based data anomaly detection method using multiple information for structural health monitoring. \textit{Structural Control and Health Monitoring}, 2019, 26.1: e2296.
\bibitem{cnn3}STAAR, Benjamin; LÜTJEN, Michael; FREITAG, Michael. Anomaly detection with convolutional neural networks for industrial surface inspection. \textit{Procedia CIRP}, 2019, 79: 484-489.
\bibitem{cnn4}BOZKURT, Baris; GERMANAKIS, Ioannis; STYLIANOU, Yannis. A study of time-frequency features for CNN-based automatic heart sound classification for pathology detection. \textit{Computers in biology and medicine}, 2018, 100: 132-143.
\bibitem{cnn5}SABOKROU, Mohammad, et al. Adversarially learned one-class classifier for novelty detection. In: \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition}. 2018. p. 3379-3388.
\bibitem{cnn6}YAHAYA, Salisu Wada; LOTFI, Ahmad; MAHMUD, Mufti. A consensus novelty detection ensemble approach for anomaly detection in activities of daily living. \textit{Applied Soft Computing}, 2019, 83: 105613.
\bibitem{cnn7}GHOSAL, Tirthankar, et al. Novelty goes deep. A deep neural solution to document level novelty detection. In: \textit{Proceedings of the 27th International Conference on Computational Linguistics}. 2018. p. 2802-2813.
\bibitem{cnn8}OZA, Poojan; PATEL, Vishal M. One-class convolutional neural network. \textit{IEEE Signal Processing Letters}, 2018, 26.2: 277-281.
\bibitem{cnn9}SCHULTHEISS, Alexander, et al. Finding the unknown: Novelty detection with extreme value signatures of deep neural activations. In: \textit{German Conference on Pattern Recognition}. Springer, Cham, 2017. p. 226-238.

\bibitem{hmm1}MACDONALD, Iain L.; ZUCCHINI, Walter. \textit{Hidden Markov and other models for discrete-valued time series}. CRC Press, 1997.
\bibitem{hmm2}BHAR, Ramaprasad; HAMORI, Shigeyuki. \textit{Hidden Markov models: applications to financial economics}. Springer Science \& Business Media, 2004.
\bibitem{hmm3}FINK, Gernot A. \textit{Markov models for pattern recognition: from theory to applications}. Springer Science \& Business Media, 2014.
\bibitem{hmm4}GALES, Mark; YOUNG, Steve. \textit{The application of hidden Markov models in speech recognition}. Now Publishers Inc, 2008.
\bibitem{hmm5}KOSKI, Timo. \textit{Hidden Markov models for bioinformatics}. Springer Science \& Business Media, 2001.
\bibitem{hmm6}WARRENDER, Christina; FORREST, Stephanie; PEARLMUTTER, Barak. Detecting intrusions using system calls: Alternative data models. In: \textit{Proceedings of the 1999 IEEE symposium on security and privacy (Cat. No. 99CB36344)}. IEEE, 1999. p. 133-145.
\bibitem{hmm7}ZOHREVAND, Zahra, et al. Hidden Markov based anomaly detection for water supply systems. In: \textit{2016 IEEE International Conference on Big Data (Big Data)}. IEEE, 2016. p. 1551-1560.
\bibitem{hmm8}YU, Jianbo. Adaptive hidden Markov model-based online learning framework for bearing faulty detection and performance degradation monitoring. \textit{Mechanical Systems and Signal Processing}, 2017, 83: 149-162.
\bibitem{hmm9}SCHMIDT, Stephan; HEYNS, Philippus Stephanus; DE VILLIERS, Johan Pieter. A novelty detection diagnostic methodology for gearboxes operating under fluctuating operating conditions using probabilistic techniques. \textit{Mechanical Systems and Signal Processing}, 2018, 100: 152-166.
\bibitem{pca}ABDI, Hervé; WILLIAMS, Lynne J. Principal component analysis. \textit{Wiley interdisciplinary reviews: computational statistics}, 2010, 2.4: 433-459.
\bibitem{hmm10}LI, Jinbo; PEDRYCZ, Witold; JAMAL, Iqbal. Multivariate time series anomaly detection: A framework of Hidden Markov Models. \textit{Applied Soft Computing}, 2017, 60: 229-240.
\bibitem{hmm11}TOLOUE, Kamran Fartash; JAHAN, Majid Vafaei. Anomalous behavior detection of marine vessels based on Hidden Markov Model. In: \textit{2018 6th Iranian Joint Congress on Fuzzy and Intelligent Systems (CFIS)}. IEEE, 2018. p. 10-12.
\bibitem{hmm12}ZHAO, Jingling, et al. Software Abnormal Behavior Detection Based on Hidden Markov Model. In: \textit{International Conference on Innovative Mobile and Internet Services in Ubiquitous Computing}. Springer, Cham, 2017. p. 929-940.
\bibitem{hmm13}LI, Peng, et al. Application of hidden Markov model in SQL injection detection. In: \textit{2017 IEEE 41st Annual Computer Software and Applications Conference (COMPSAC)}. IEEE, 2017. p. 578-583.
\bibitem{hmm14}ROBINSON, William N.; ARIA, Andrea. Sequential fraud detection for prepaid cards using hidden Markov model divergence. \textit{Expert Systems with Applications}, 2018, 91: 235-251.

\bibitem{expo}SHARMA, Anish and ANDREWS, Rebecca. Managing-Exponential-Data-Growth-and-Application-Modernization. \textit{IBM} [online]. 11 November 1999. [cit. 11.8.2020]. Dostupné z: https://www.ibm.com/cloud/blog/managing-exponential-data-growth-and-application-modernization 
\end{thebibliography}



\clearpage
\chapter*{Příloha}
\markboth{Příloha}{}
\addcontentsline{toc}{chapter}{Příloha}
\begin{appendices}

\end{appendices}



\end{document}